\documentclass[]{article}

\usepackage{listings}
\usepackage{hyperref}
\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black
}

%opening
\title{CSE150 -- Project 3\\ Networks and Distributed Systems}
\author{Group Thimbles}

\begin{document}

\maketitle
\tableofcontents

\section{Data structures}

\subsection{Socket object}
\paragraph{Overview}
The socket object keeps track of all state associated with socket operations, including the TCP state machine and the mapped ports, as well as the associated \texttt{FileObject} file descriptor.


\subsubsection{Send and receive buffers}

\subsubsection{TCP state machine}

The socket object contains a TCP state machine implementing the seven TCP states and the transitions between them, which occur based on syscalls, timer events or the receipt of TCP-flagged packets, such as \texttt{ACK} or \texttt{STP}.

\subsection{Client object}

\subsection{Sliding window}

\subsection{Chat}

\section{Networking syscalls}
\subsection{connect()}

\paragraph{Overview}
The \texttt{connect()} syscall takes as arguments the host address and the port to which a connection will be opened. It begins by disabling interrupts and creating a new socket object s, which is assigned a valid, free port in the range (0,127). Next, the \texttt{state} member is set to \texttt{SYN\_SENT} and a \texttt{SYN} packet is sent out on the connection. 

At this point, the socket blocks its handler thread until it receives a SYN/ACK interrupt from the network layer. Alternatively, a timeout from the alarm clock can break the code (how?). Once the \texttt{SYN/ACK} has arrived, \texttt{s.state} is set to \texttt{ESTABLISHED} and interrupts are re-enabled. At this point, the connection has been successfully opened and the function returns the integer value of socket's associated file descriptor. \\






\noindent Pseudocode:

\begin{lstlisting}
int connect(host, port){
	create a new socket s
	assign s a free port in (0,127)
	s.state = SYN_SENT
	send SYN packet
	block until SYN/ACK recv'd // timeout breaks this?
	s.state = ESTABLISHED
	return s.fileDescriptor
}
\end{lstlisting}

\subsection{accept()}
\paragraph{Overview}
The \texttt{accept()} syscall takes as arguments a port on which to accept a connection and is serviced by the receive thread. It begins by checking for a waiting connection on the supplied port. If there is a waiting connection, a new \texttt{socket} object is created and assigned to that port. Otherwise, -1 is immediately returned.

Next, the TCP state of the socket is set to \texttt{ESTABLISHED} and a \texttt{SYN/ACK} packet is sent out. Lastly, the method returns the socket's associated file descriptor.\\

\noindent Pseudocode:

\begin{lstlisting}
int acccept(port){
	if there are connections waiting on port
		create a new socket s, assign it that port
	else return -1
	s.state = ESTABLISHED
	send SYN/ACK
	return s.fileDescriptor
}
\end{lstlisting}

\subsection{read()}
\paragraph{Overview}
The \texttt{read()} syscall extends the implementation from phase 2 for the case in which it encounters a \texttt{socket} object. As in the previous implementation, \texttt{read()} takes as input the file descriptor, a buffer into which to place the read data and a number of bytes of data to read.

If the socket is open, \texttt{read()} will read up to the number of requested bytes, assuming that there are that many in the buffer. If there are fewer than requested then the entire buffer is retrieved and the number of bytes read is returned.

If the socket is closed and still has data, then read will work in the same way as before, with the difference that it will delete the socket. 

\begin{lstlisting}
int read(fileDescriptor, buffer, count){
	...
	// for a socket
	if (s.isOpen){
		read count bytes
		return bytes successfully read
	} else {
		if (socket isn't empty){
			read count bytes
			if (socket is empty)
				delete socket
			return bytes successfully read	
		}	
	}
}
\end{lstlisting}

\subsection{write()}
\paragraph{Overview}
The \texttt{write()} syscall extends the implementation from phase 2 with special handling for sockets. It attempts to write the supplied number of bytes from the supplied buffer into the specified socket, returning -1 on failure. It succeeds if the connection is already established and the write (outgoing) buffer has room for the entire message. Alternatively, the implementation could write up to the end of the outgoing buffer and return the number of bytes written.\\

\noindent Pseudocode: 

\begin{lstlisting}
int write(fileDescriptor, buffer, count){
	...
	// no partial writes
	if socket is ESTABLISHED and the write buffer has room
		write to the outgoing buffer
		//post office handles packet creation
	...
}
\end{lstlisting}

\subsection{open()}

\subsection{close()}
\paragraph{Overview}
The \texttt{close()} syscall extends the existing implementation to handle sockets and is serviced by the \texttt{receive} thread. It begins by checking the state of the socket; if the socket is in state \texttt{ESTABLISHED} and the outgoing window is empty, it will send a \texttt{FIN} packet to the other end and set the socket state to \texttt{CLOSING}. 

If there are still pending packets in the outgoing window, \texttt{close()} sends an \texttt{STP} packet to the other end and sets the socket state to \texttt{STP\_SENT}. Attempting to close a socket that is not in \texttt{ESTABLISHED} returns -1.

\pagebreak

\noindent Pseudocode:

\begin{lstlisting}
int close(fileDescriptor) {
	if s.state == ESTABLISHED {
		if outgoing window is empty {
			send FIN;
			s.state = CLOSING;
		} else {
			send STP;
			s.state = STP_SENT;
			return 0;
		}
	} else return -1;
}
\end{lstlisting}


\section{Threads}
\paragraph{Overview}
There are three primary threads responsible for the handling of network communication, \texttt{send}, \texttt{receive} and \texttt{timeout}.

\subsection{Send thread}
This \texttt{send} thread iterates through all of the existing socket's outgoing buffers and sends the data out as expected.
\subsection{Receive thread}
The \texttt{receive} thread iterates through all of the existing socket's incoming buffers and passes the data up.
\subsection{Timeout thread}
This thread works like waitUntil, where it loops through the existing sockets and checks for any that have lived past their timeout value. If they have, it closes that socket.

\section{Chat}
\subsection{Client}
\paragraph{Overview} The chat client maintains three buffers, \texttt{outgoingMessage}, \\ \texttt{incomingMessage} and \texttt{temp} and has a socket connection to the server on (server-incoming) port 15. Each of the buffers has size equal to the predefined maximum character count for a line.

\texttt{outgoingMessage} is used to store and buffer user-typed messages, \\ \texttt{incomingMessage} buffers messages from the server and \texttt{temp} is used as a swap buffer.

The client itself runs in an indefinite loop, until it receives a '.' character as input, at which point it closes its socket connection and returns exit code 0 for success. 

In each loop iteration, the client first reads pending user input into \\ \texttt{outgoingMessage}, copying it past any existing data in the buffer. It then checks that there is user input to be handled, and if there is, loops through \texttt{outgoingMessage}, copying data up the newline character into \texttt{temp} and moving the remaining data to the beginning of the buffer. \texttt{temp} is then written to the socket. This continues until there are no more newline characters in \texttt{outgoingMessage}. In this way, users can send multiple lines of text (separated by newline characters) with their order preserved (i.e. the server will always receive them in the order they are typed).

If there is no user input pending, then the client performs a \texttt{read} operation on the socket, filling \texttt{incomingMessage} with data starting from the end of the existing data, up to the end of the buffer. After the read completes, the client loops through \texttt{incomingMessage}, copying data into \texttt{temp} up to each \textbackslash n character, moving the data in \texttt{incomingMessage} to the beginning of the buffer and printing the contents of \texttt{temp} to the screen. \\

\pagebreak

\noindent Pseudocode:

\begin{lstlisting}
int main(serverAddress){
  create incomingMessage, outgoingMessage and temp buffers
  create a socket connected to destination (serverAddress,15)
  while (true){
    // fill outgoingMessage up to the end
    bytesToRead = MAX_CHAR_COUNT - len(outgoingMessage)
	read bytesToRead bytes into outgoingMessage from user input
	if outgoingMessage has data {
	  // loop until outgoingMessage is empty
  	  while outgoingMessage has a '\n' {
	  copy characters up to '\n' into temp
	  if temp is '.'
  	  close s
	  return 0 for success
	  write temp to the socket
	  }
	} else {
	  // fill incomingMessage up to the end
	  bytesToRead = MAX_CHAR_COUNT - len(incomingMessage)
	  read bytesToRead bytes from socket into incomingMessage
	  if incomingMessage has data {
	    // loop until incomingMessage is empty
	    while incomingMessage has a '\n' {
	      copy characters up to '\n' into temp
	      'shift' everything past '\n' to the beginning
	      print contents of temp to screen
	    }
	  }
	} 	
}
\end{lstlisting}
\subsection{Server}
\paragraph{Overview}
The chat server maintains a single buffer, \texttt{temp}, of size equal to maximum number of allowed characters and runs in an indefinite loop. It immediately terminates upon input on \texttt{stdin}, returning 0 for success.

Within the loop body, the server calls \texttt{accept()} on the supplied port and allocates a \texttt{client} object if there is a pending connection. It then iterates through its list of clients, reading from the clients' associated socket into the temporary buffer and copying that buffer into the client's \texttt{outgoingMessage} buffer. The server then iterates through \texttt{client.outgoingMessage}, copying and moving over newline-delimited character arrays (into \texttt{temp}) and performing \texttt{write()} operations on every client's socket.

\pagebreak

\noindent Pseudocode:

\begin{lstlisting}
int main(listenPort){
  create temp buffer of size MAX_CHAR_COUNT
  while (true){
    // if stdin then return 0 for success
    // accept a connection on listenPort and allocate a client
    // client object holds socket s
    for each client:
      read from client.s into temp
      copy temp into client.outgoingMessage
      while (client.outgoingMessage has '\n'){
        copy chars up to '\n' into temp
        'shift' everything past '\n' to the beginning
        for each client:
          write temp to client.s
      }
  }
}	
\end{lstlisting}

\section{Test cases}
\subsection{connect()}
\begin{itemize}
	\item Attempt to open a connection to a node that doesn't exist
	\subitem Check that connect() blocks
	\item Open a connection to an existing node
	\subitem Check that connect() returns
	\item Close an already-open connection
	\subitem Verify that socket is closed on both sides
	\item Open multiple connections to the same receiving port
	\subitem Check that they all send/receive data
	\item Open a connection, close it and re-open it
\end{itemize}
\subsection{accept()}
\begin{itemize}
	\item Accept a waiting connection
	\item Accept multiple waiting connections on the same port
	\item Accept multiple waiting connections to different ports
	\item Return from accept() on a port that doesn't have a connection waiting
\end{itemize}
\subsection{close()}
\begin{itemize}
	\item Close a connection that doesn't exist
	\item Close a connection that exists
	\subitem Check that it's actually closed
	\item Close a connection as the sender and as the receiver
	\subitem Check 
	\item Close a connection twice in a row
	\subitem Check that errors don't occur
	\item Close a connection with a lot of data waiting
	\subitem Check that sender 
\end{itemize}
\subsection{read()}
\begin{itemize}
	\item Read from an open socket
	\subitem Check that it doesn't block
	\item Read from a closed socket with data remaining
	\subitem Check that it doesn't block
	\subitem Check that data is returned
	\subitem Return value?
	\item Read from an empty socket
	\subitem Check that it returns 0 w/o blocking
\end{itemize}
\subsection{write()}
\begin{itemize}
	\item Write to an open socket
	\item Write to a closed socket
	\subitem Check that it returns -1 immediately
\end{itemize}
\subsection{send()}
\subsection{receive()}
\subsection{Connection}
\begin{itemize}
	\item Write to a connection and read from that connection on the receiving node
	\subitem Check that output matches input
	\subitem Check that window size never exceeds 16 packets (how?)
	\item Write to both sides of a connection simultaneously
	\item Read from both sides of a connection simultaneously
	\item Write/read to/from a connection with 10\% loss rate
	\subitem Check that output matches input
	\item Write a large amount of data into an open connection and read it on the receiving node
	\subitem Check that output matches input
\end{itemize}
\subsection{Chat client/server}
\begin{itemize}
	\item Broadcast a message to all clients
	\subitem Check that it works for small numbers and large numbers
	\subitem Check that output matches input for every client
	\subitem Check that sending a message doesn't interleave with receiving a message
	\item Connect a client, then drop the client
	\item Connect a client, drop the client, then reconnect that same client
	\item Connect multiple clients
	\item Drop multiple clients
	\item Reconnect multiple clients
	\item Broadcast multiple messages
	\subitem Check that the order of the messages is preserved for a given sender
	\item Quit the client
	\subitem Check that client exists gracefully (how?) after getting a single '.'
	\item Quit the server
	\subitem Check that the server exists upon receiving anything on stdin
	\subitem Check that all clients get a message about server closing
	
\end{itemize}
\end{document}
